---
title: "NYCabs CA-MCA-Clustering"
author: "L?dia Montero"
date: "13 November 2020"
output: 
  html_document: 
    toc: true
    toc_depth: 3
    number_sections: true
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Description http://www.nyc.gov/html/tlc/html/about/trip_record_data.shtml
Data Dictionary - SHL Trip Records -This data dictionary describes SHL trip data in visit http://www.nyc.gov/html/tlc/html/about/trip_record_data.shtml.    

  - VendorID	A code indicating the LPEP provider that provided the record.      1= Creative Mobile Technologies, LLC; 2= VeriFone Inc.   
  - lpep_pickup_datetime	The date and time when the meter was engaged.    
  - lpep_dropoff_datetime	The date and time when the meter was disengaged.     
  - Passenger_count	The number of passengers in the vehicle. This is a driver-entered value.    
  -  Trip_distance 	The elapsed trip distance in miles reported by the taximeter.   - Pickup_longitude	 Longitude where the meter was engaged.   
  - Pickup_latitude	Latitude where the meter was engaged.   
  - RateCodeID	The final rate code in effect at the end of the trip.  1= Standard rate  2=JFK 3=Newark 4=Nassau or Westchester 5=Negotiated fare 6=Group ride   
  - Store_and_fwd_flag	This flag indicates whether the trip record was held in vehicle memory before sending to the vendor, aka "store and forward," because the vehicle did not have a connection to the server: Y= store and forward trip  N= not a store and forward trip   
  - Dropoff_longitude	Longitude where the meter was timed off.   
  - Dropoff_ latitude	Latitude where the meter was timed off.   
  - Payment_type	A numeric code signifying how the passenger paid for the trip.  1= Credit card 2= Cash 3= No charge 4= Dispute 5= Unknown 6= Voided trip   
  - Fare_amount	The time-and-distance fare calculated by the meter.   
  - Extra	 Miscellaneous extras and surcharges.  Currently, this only includes the $0.50 and $1 rush hour and overnight charges. 
  - MTA_tax	 $0.50 MTA tax that is automatically triggered based on the metered rate in use.    - Improvement_surcharge	$0.30 improvement surcharge assessed on hailed trips at the flag   drop. The improvement surcharge began being levied in 2015.   
  - Tip_amount	 This field is automatically populated for credit card tips. Cash tips are not included.   
  - Tolls_amount	Total amount of all tolls paid in trip.    
  - Total_amount	The total amount charged to passengers. Does not include cash tips.   
  - Trip_type	A code indicating whether the trip was a street-hail or a dispatch that is automatically assigned based on the metered rate in use but can be altered by the driver. 
1= Street-hail 2= Dispatch  


```{r}
# Load Required Packages: to be increased over the course
rm(list=ls())
requiredPackages <- c("effects","FactoMineR","car", "factoextra","RColorBrewer","ggplot2","missMDA","mvoutlier","dplyr","ggmap","ggthemes","knitr","MVA")
missingPackages <- requiredPackages[!(requiredPackages %in% installed.packages()[,"Package"])]

if(length(missingPackages)) install.packages(missingPackages)
lapply(requiredPackages, require, character.only = TRUE)

# Useful function
calcQ <- function(x) {
  s.x <- summary(x)
  iqr<-s.x[5]-s.x[2]
  list(souti=s.x[2]-3*iqr, mouti=s.x[2]-1.5*iqr, min=s.x[1], q1=s.x[2], q2=s.x[3],        q3=s.x[5], max=s.x[6], mouts=s.x[5]+1.5*iqr, souts=s.x[5]+3*iqr ) }
```

# Load your sample after data cleaning and validation (Deliverable 1)


```{r}
setwd("C:/Users/lmontero/Dropbox/DOCENCIA/FIB-ADEI/PRACTICA/NYCABS/LABS") # Choose your own directory
file_path <- "C:/Users/lmontero/Dropbox/DOCENCIA/FIB-ADEI/PRACTICA/NYCABS/LABS/" 
load(paste0(file_path,"MyTaxi5000Clean.RData"))
names(df)
summary(df)
vars_con
vars_dis
vars_res
```

# CA: f.cost (discretization of Total_amount) vs f.hour and f.tt and period ...

```{r}
par(mfrow=c(1,1))

tt<-table(df[,c("f.cost","period")])
tt
res.ca<-CA(tt)
#lines(res.ca$row$coord[,1],res.ca$row$coord[,2],lwd=2,col="blue")
chisq.test(tt)

names(res.ca)
fviz_eig(res.ca)
summary(res.ca,dig=2)
fviz_ca_biplot(res.ca,repel=TRUE)+theme_bw()

tt<-table(df[,c("f.cost","f.tt")])
tt
res.ca<-CA(tt)
#lines(res.ca$row$coord[,1],res.ca$row$coord[,2],lwd=2,col="blue")
summary(res.ca)
chisq.test(tt)

tt<-table(df[,c("f.cost","f.hour")])
tt
res.ca<-CA(tt)
ellipseCA(res.ca,ellipse="row")
#lines(res.ca$row$coord[,1],res.ca$row$coord[,2],lwd=2,col="blue")
#lines(res.ca$col$coord[,1],res.ca$col$coord[,2],lwd=2,col="red")
summary(res.ca)

# Traditional analysis by contingency tables
prop.table(tt,1)
prop.table(tt,2)
chisq.test(tt)
```

# MCA using multivariant

```{r}
par(mfrow=c(1,1))
#llvout<-which(df$mvout==TRUE);length(llvout)

#res.mca<-MCA(df[,c(vars_dis,"Total_amount")],quali.sup=c(5,14),quanti.sup=19,ind.sup=llvout)
res.mca<-MCA(df[,c(vars_dis,"Total_amount")],quali.sup=c(5,14),quanti.sup=19)
summary(res.mca,nbelements=50)
mean(res.mca$eig[,1]) # Indica fins dim 23 o criteri Inercia explicada 80% ?s 30

# Individual Representation
plot.MCA(res.mca,choix=c("ind"),cex=0.8)
plot.MCA(res.mca,choix=c("ind"),invisible=c("ind"),cex=0.8)

# Representation of categories
plot.MCA(res.mca,choix=c("ind"),invisible=c("ind"),axes=c(1,2))
#lines(res.mca$var$coord[11:14,1],res.mca$var$coord[11:14,2],lwd=1,col="black") # Speed
#lines(res.mca$quali.sup$coord[3:6,1],res.mca$quali.sup$coord[3:6,2],lwd=2,col="darkgreen")
names(res.mca)
res.mca$var
plot.MCA(res.mca,choix=c("ind"),invisible=c("ind"),axes=c(3,4))
plot.MCA(res.mca,choix=c("var"),axes=c(3,4))

# Use modern ggplot facilities
names(res.mca)
fviz_eig(res.mca)
fviz_cos2(res.mca, choice = "var", axes = 1:2)+theme_bw()
fviz_contrib(res.mca, choice = "var", axes = 1:2)+theme_bw()
fviz_mca_var(res.mca, col.var="contrib",repel=TRUE,labelsize = 2)+
    scale_color_gradient2(low="green", mid="blue", 
    high="red", midpoint=0.75)+theme_bw()

fviz_mca_biplot(res.mca, invisible="ind",axes=1:2,repel=FALSE,labelsize = 2)+theme_bw()
#fviz_mca_biplot(res.mca, axes=1:2,repel=TRUE,arrows=c(TRUE,FALSE))+theme_bw()
```

# Synthesis through HCPC: Hierarchical Clustering

```{r}
###
### Clustering the individuals
### Before, hou have to perform a PCA with the number of axes 
### that you have decided to take into account (indicated through ncp=)
?HCPC
#res.mca<-MCA(df[,c(vars_dis,"Total_amount")],quali.sup=c(5,14),quanti.sup=19,ncp=35,ind.sup=llvout)
res.mca<-MCA(df[,c(vars_dis,"Total_amount")],quali.sup=c(5,14),quanti.sup=19,ncp=35)
res.hcpc<-HCPC(res.mca,nb.clust=8,order=TRUE)
names(res.hcpc)

names(res.hcpc$call$t)
res.hcpc$call$t$within[1:8]
(res.hcpc$call$t$within[1]-res.hcpc$call$t$within[1:15])/res.hcpc$call$t$within[1]

df$claHP<-9
df[row.names(res.hcpc$data.clust),"claHP"]<-res.hcpc$data.clust$clust
df$claHP<-factor(df$claHP)
summary(res.hcpc$data.clust$clust)
table(df$claHP)

### Interpret clustering results
summary(res.hcpc$data.clust$clust)

#Block A: res.hcpc$desc.var
#Block B: res.hcpc$desc.axes
#Block C: res.hcpc$desc.ind


### desc.var ###
### A. The description of the clusters by the variables ###
names(res.hcpc$desc.var)
res.hcpc$desc.var
### desc.var$test.chi2 ###
### A.1. The categorical variables which characterizes the clusters ###
res.hcpc$desc.var$test.chi2

### desc.var$category ###
### A.2. The description of each cluster by the categories ##
res.hcpc$desc.var$category

### desc.var$quanti.var ###
### A.3. The quantitative variables which characterizes the clusters ###
res.hcpc$desc.var$quanti.var

### desc.var$quanti ###
### A.4. The description of each cluster by the quantitative variables ###
res.hcpc$desc.var$quanti

### desc.axes ###
### B. The description of the clusters by the axes ###
names(res.hcpc$desc.axes)
res.hcpc$desc.axes
res.hcpc$desc.axes$quanti.var
res.hcpc$desc.axes$quanti

### desc.ind ###
### C. The description of the clusters by the individuals ###
names(res.hcpc$desc.ind)
res.hcpc$desc.ind$para
res.hcpc$desc.ind$dist

#### Characteristic individuals
para1<-which(rownames(res.mca$ind$coord)%in%names(res.hcpc$desc.ind$para[[1]]))
para2<-which(rownames(res.mca$ind$coord)%in%names(res.hcpc$desc.ind$para[[2]]))
para3<-which(rownames(res.mca$ind$coord)%in%names(res.hcpc$desc.ind$para[[3]]))
# to be completed...

#plot(res.mca,label="none",invisible=c("quali","ind.sup"))
#points(res.mca$ind$coord[para1,1],res.mca$ind$coord[para1,2],col="magenta",cex=2,pch=16)

# Dendrogram
fviz_dend(res.hcpc, show_labels = FALSE)
# Individuals factor map
fviz_cluster(res.hcpc, geom = "point", main = "Factor map")

res.mca<-MCA(df[,c(vars_dis,"Total_amount","claHP")],quali.sup=c(5,14,20),quanti.sup=19,ncp=35)
fviz_mca_ind(res.mca, 
             label = "none", invisible=c("ind.sup"),# hide individual labels
             habillage = "claHP", # color by groups 
             addEllipses = TRUE, ellipse.type = "confidence",
             ggtheme = theme_bw()) 
```

# Kmeans: Partitioning in k=8 groups
```{r}
res.mca<-MCA(df[,c(vars_dis,"Total_amount")],quali.sup=c(5,14),quanti.sup=19,ncp=29)
summary(res.mca)
ppcc<-res.mca$ind$coord[,1:29]
dim(ppcc)
kc<-kmeans(dist(ppcc),8, iter.max = 30, trace=T)

df[names(kc$cluster),"claKM"]<-kc$cluster
kc$betweenss/kc$totss

```

