---
title: "Deliverable I"
author: 
- name: Jesús Molina Roldán
- name: Carles Capilla Cànovas
date: \today
output:
  word_document:
    toc: no
    toc_depth: '4'
  md_document:
  html_document:
    toc: no
    toc_depth: '4'
  pdf_document:
    number_sections: yes
    toc: yes
    toc_depth: '4'
geometry: left=1.9cm,right=1.9cm,top=1.25cm,bottom=1.52cm
fontsize: 18pt
subtitle: 'ADEI'
classoption: a4paper
editor_options: 
  chunk_output_type: console
---
## Data Description

This data dictionary describes SHL trip data. For dictionaries describing yellow taxi and FHV data, please visit http://www.nyc.gov/html/tlc/html/about/trip_record_data.shtml.

  - VendorID	A code indicating the LPEP provider that provided the record.      1= Creative Mobile Technologies, LLC; 2= VeriFone Inc.   
  - lpep_pickup_datetime	The date and time when the meter was engaged.    
  - lpep_dropoff_datetime	The date and time when the meter was disengaged.     
  - Passenger_count	The number of passengers in the vehicle. This is a driver-entered value.    -  Trip_distance 	The elapsed trip distance in miles reported by the taximeter.   
  - Pickup_longitude	 Longitude where the meter was engaged.   
  - Pickup_latitude	Latitude where the meter was engaged.   
  - RateCodeID	The final rate code in effect at the end of the trip.  1= Standard rate  2=JFK 3=Newark 4=Nassau or Westchester 5=Negotiated fare 6=Group ride   
  - Store_and_fwd_flag	This flag indicates whether the trip record was held in vehicle memory before sending to the vendor, aka "store and forward," because the vehicle did not have a connection to the server: Y= store and forward trip  N= not a store and forward trip   
  - Dropoff_longitude	Longitude where the meter was timed off.   
  - Dropoff_ latitude	Latitude where the meter was timed off.   
  - Payment_type	A numeric code signifying how the passenger paid for the trip.  1= Credit card 2= Cash 3= No charge 4= Dispute 5= Unknown 6= Voided trip   
  - Fare_amount	The time-and-distance fare calculated by the meter.   
  - Extra	 Miscellaneous extras and surcharges.  Currently, this only includes the $0.50 and $1 rush hour and overnight charges. 
  - MTA_tax	 $0.50 MTA tax that is automatically triggered based on the metered rate in use.    - Improvement_surcharge	$0.30 improvement surcharge assessed on hailed trips at the flag   drop. The improvement surcharge began being levied in 2015.   
  - Tip_amount	 This field is automatically populated for credit card tips. Cash tips are not included.   
  - Tolls_amount	Total amount of all tolls paid in trip.    
  - Total_amount	The total amount charged to passengers. Does not include cash tips.   
  - Trip_type	A code indicating whether the trip was a street-hail or a dispatch that is automatically assigned based on the metered rate in use but can be altered by the driver. 
1= Street-hail 2= Dispatch  

# Load Required Packages: to be increased over the course

```{r echo=T, results='hide', warning=FALSE}
options(contrasts=c("contr.treatment","contr.treatment"))

requiredPackages <- c("missMDA","chemometrics","mvoutlier","effects","FactoMineR","car", "factoextra","RColorBrewer","ggplot2","dplyr","ggmap","ggthemes","knitr")
missingPackages <- requiredPackages[!(requiredPackages %in% installed.packages()[,"Package"])]

if(length(missingPackages)) install.packages(missingPackages)
lapply(requiredPackages, require, character.only = TRUE)

```



```{r, echo=FALSE, results='FALSE'}
# Clear plots
if(!is.null(dev.list())) dev.off()

# Clean workspace
rm(list=ls())

#setwd("C:/Users/carle/Desktop/ADEI2")
#filepath<-"C:/Users/carle/Desktop/ADEI2/"
# green_tripdata_2016-01

setwd("~/Desktop/NYCABS")
filepath<-"~/Desktop/NYCABS/"
# green_tripdata_2016-01

df2<-read.table(paste0(filepath,"green_tripdata_2016-01.csv"),header=T, sep=",")
#Reload data
dim(df2)
names(df2)
head(df2)
```

## Select a sample of 5000 records

```{r, echo=TRUE, results=FALSE}
set.seed(02041997)
sam<-as.vector(sort(sample(1:nrow(df2),5000)))
df<-df2[sam,]
summary(df)
```

## Some useful functions

```{r, echo=TRUE, results=FALSE}
calcQ <- function(x) {
  s.x <- summary(x)
  iqr<-s.x[5]-s.x[2]
  list(souti=s.x[2]-3*iqr, mouti=s.x[2]-1.5*iqr, min=s.x[1], q1=s.x[2], q2=s.x[3], 
       q3=s.x[5], max=s.x[6], mouts=s.x[5]+1.5*iqr, souts=s.x[5]+3*iqr ) }

countNA <- function(x) {
  mis_x <- NULL
  for (j in 1:ncol(x)) {mis_x[j] <- sum(is.na(x[,j])) }
  mis_x <- as.data.frame(mis_x)
  rownames(mis_x) <- names(x)
  mis_i <- rep(0,nrow(x))
  for (j in 1:ncol(x)) {mis_i <- mis_i + as.numeric(is.na(x[,j])) }
  list(mis_col=mis_x,mis_ind=mis_i) }

countX <- function(x,X) {
  n_x <- NULL
  for (j in 1:ncol(x)) {n_x[j] <- sum(x[,j]==X) }
  n_x <- as.data.frame(n_x)
  rownames(n_x) <- names(x)
  nx_i <- rep(0,nrow(x))
  for (j in 1:ncol(x)) {nx_i <- nx_i + as.numeric(x[,j]==X) }
  list(nx_col=n_x,nx_ind=nx_i) }

EDA_Univariant <- function(x,title,bins){
  listOfColors<-rainbow(39)
  if(missing(title)) {
        title<-""
  }
  if(missing(bins)){
      bins <- 100
  }
  x <- data.frame(x)
  print(summary(x))
  par(mfrow=c(NCOL(x),2))
  for (i in 1:NCOL(x)){
      if(NCOL(x)>1){
        title <-names(x)[i]
      }
      if (sapply(x, class)[i] == "numeric" || sapply(x, class)[i] == "integer"){
        print(sd(x[,i]))
        hist(x[,i], main = paste("Histogram", title), freq = TRUE, xlab="", col =  listOfColors, breaks = bins)
        boxplot(x[,i], main = paste("Boxplot", title))
        var_out<-calcQ(df$espeed)
        abline(h=var_out$souts,col="red")
        abline(h=var_out$souti,col="blue")
      }else{
        barplot(table(as.factor(x[,i])), main = paste("Barplot", title), col =  listOfColors)
        pie(table(as.factor(x[,i])), main = paste("Pie", title), col =  listOfColors)
      }
  }
  par(mfrow=c(1,1))
}

```

## Factors: Levels coding

Now codify properly factors and remove non-informative variables
```{r, echo=FALSE}
summary(df)
names(df)
df$VendorID<-factor(df$VendorID,labels=c("Mobile","VeriFone"))
levels(df$VendorID)<-paste0("f.Vendor-",levels(df$VendorID))
summary(df$VendorID)

df$Trip_type<-factor(df$Trip_type,labels=c("Street-Hail","Dispatch"))
levels(df$Trip_type)<-paste0("f.TripType-",levels(df$Trip_type))
summary(df$Trip_type)
nlevels(df$Trip_type)

df$Payment_type<-factor(df$Payment_type,labels=c("Credit card","Cash","No charge","Dispute"))
levels(df$Payment_type)<-paste0("f.PayType-",levels(df$Payment_type))
summary(df$Payment_type)
nlevels(df$Payment_type)


colnames(df)[which(names(df) == "Lpep_dropoff_datetime")] <- "lpep_dropoff_datetime"


summary(df$Store_and_fwd_flag)
df$Store_and_fwd_flag <- as.factor(df$Store_and_fwd_flag == "Y")

summary(df$improvement_surcharge)
colnames(df)[which(names(df) == "improvement_surcharge")] <- "Improvement_surcharge"

```


```{r}
# Initialization of counts for missings, outliers and errors. All numerical variables have to be checked before

#######################################################
imis<-rep(0,nrow(df))  # rows - trips
jmis<-rep(0,2*ncol(df))  # columns - variables
######################################################
jmis <- colSums(is.na(df))
imis <- rowSums(is.na(df))
jmis
imis
#######################################################
iouts<-rep(0,nrow(df))  # rows outliers - trips
jouts<-rep(0,2*ncol(df))  # columns outliers - variables
######################################################

#######################################################
ierrs<-rep(0,nrow(df))  # rows errors - trips
jerrs<-rep(0,2*ncol(df))  # columns errors - variables
######################################################


```


## Univariant outliers and errors

```{r}
####   Variable RateCodeID. Consider Recoding to fewer levels
summary(df$RateCodeID)
df$RateCodeID<-factor(df$RateCodeID)
barplot(table(df$RateCodeID))

# It is a categorical(=factor) variable   NO PROBLEM but not any interest
df$RateCodeID <- df$RateCodeID != 1
df$RateCodeID <- factor(df$RateCodeID, labels=c("Standard rate","Others"))
summary(df$RateCodeID)
barplot(table(df$RateCodeID))

#Variable Passenger_count

hist(df$Passenger_count, main="Histogram of Passenger_count")
boxplot(df$Passenger_count, main="Boxplot of Passenger_count")
summary(df$Passenger_count)

# errors
l <- which(df$Passenger_count<0)
if (length(l)>0) {
  ierrs[l]<-ierrs[l]+1
  jerrs["Passenger_count"]<-length(l)
}
df[l,"Passenger_count"]<-NA 


hist(df$Passenger_count, main="Histogram of Passenger_count")
boxplot(df$Passenger_count, main="Boxplot of Passenger_count")
summary(df$Passenger_count)


#Variable Trip_distance
summary(df$Trip_distance)
# errors
l <- which(df$Trip_distance<0.001);length(l)
if (length(l)>0) {
  ierrs[l]<-ierrs[l]+1
  jerrs["Trip_distance"]<-length(l)
}
df[l,"Trip_distance"]<-NA 



#outliers
hist(df$Trip_distance, main="Histogram of Trip_distance")
boxplot(df$Trip_distance, main="Boxplot of Trip_distance")
var_out<-calcQ(df$Trip_distance)
abline(h=var_out$souts,col="red")
abline(h=var_out$souti,col="blue")
l<-which(df$Trip_distance>20)
iouts[l]<-iouts[l]+1
jouts["Trip_distance"]<-length(l)



### Variable Pickup_latitude
l<-which(df$Pickup_latitude ==0)
ierrs[l]<-ierrs[l]+1
jerrs["Pickup_latitude"]<-length(l)
df[l,"Pickup_latitude"]<-NA    

### Variable Pickup_longitude
l<-which(df$Pickup_longitude ==0)
ierrs[l]<-ierrs[l]+1
jerrs["Pickup_longitude"]<-length(l)
df[l,"Pickup_longitude"]<-NA    

### Variable Dropoff_latitude
l<-which(df$Dropoff_latitude ==0)
ierrs[l]<-ierrs[l]+1
jerrs["Dropoff_latitude"]<-length(l)
df[l,"Dropoff_latitude"]<-NA    

### Variable Dropoff_longitude
l<-which(df$Dropoff_longitude ==0)
ierrs[l]<-ierrs[l]+1
jerrs["Dropoff_longitude"]<-length(l)
df[l,"Dropoff_longitude"]<-NA  



####   Variable Store_and_fwd_flag
summary(df$Store_and_fwd_flag)
df$Store_and_fwd_flag<-factor(df$Store_and_fwd_flag)
barplot(table(df$Store_and_fwd_flag))
# It is a categorical(=factor) variable   NO PROBLEM but not any interest


### Variable Payment_type
summary(df$Payment_type)
frequency <- summary(df$Payment_type)
frequency <- frequency/sum(frequency)
frequency
barplot(frequency,ylim=c(0,1))
barplot(table(df$Payment_type))

l <- which(df$Payment_type != "f.PayType-Credit card" & df$Payment_type != "f.PayType-Cash" )
df$Payment_type <- as.character(df$Payment_type)
df[l,"Payment_type"] <- "f.PayType-Others"
df$Payment_type <- as.factor(df$Payment_type )
summary(df$Payment_type)
barplot(table(df$Payment_type))

## Variable Fare_amount
hist(df$Fare_amount, main="Histogram of Fare_amount")
boxplot(df$Fare_amount, main="Boxplot of Fare_amount")
summary(df$Fare_amount)

# errors
l <- which(df$Fare_amount<0)
if (length(l)>0) {
  ierrs[l]<-ierrs[l]+1
  jerrs["Fare_amount"]<-length(l)
}
df[l,"Fare_amount"]<-NA 

# Outlier detection
boxplot(df$Fare_amount)
var_out<-calcQ(df$Fare_amount)
abline(h=var_out$souts,col="red")
abline(h=var_out$souti,col="blue")
var_out$souts
l <-which(df$Fare_amount>80)
iouts[l ]<-iouts[l]+1
jouts["Fare_amount"]<-length(l)
df[l,"Fare_amount"]<-NA 

hist(df$Fare_amount, main="Histogram of Fare_amount")
Boxplot(df$Fare_amount, main="Boxplot of Fare_amount")
summary(df$Fare_amount)

## Variable Extra

hist(df$Extra, main="Histogram of Extra")
boxplot(df$Extra, main="Boxplot of Extra")
summary(df$Extra)

# errors
l <- which(df$Extra!= 0.5 || df$Extra != 1 || df$Extra != 0)
if (length(l)>0) {
  ierrs[l]<-ierrs[l]+1
  jerrs["Extra"]<-length(l)
}

df[,"f.Extra"] <- "f.Extra-0"
l <- which(df$Extra >0.4)
df[l,"f.Extra"] <- "f.Extra-0.5"
l <- which(df$Extra >0.9)
df[l,"f.Extra"] <- "f.Extra-1"
df[,"f.Extra"] <- as.factor(df[,"f.Extra"])

### Variable MTA_tax
hist(df$MTA_tax, main="Histogram of MTA_tax")
boxplot(df$MTA_tax, main="Boxplot of MTA_tax")
summary(df$MTA_tax)

# errors
l <- which(df$MTA_tax<0)
if (length(l)>0) {
  ierrs[l]<-ierrs[l]+1
  jerrs["MTA_tax"]<-length(l)
}
df[l,"MTA_tax"]<-NA 

df[,"MTA_tax"] <- as.factor(df[,"MTA_tax"] == 0.5)
barplot(table(df[,"MTA_tax"]))
summary(df[,"MTA_tax"] )

#it is not a useful information because in theory it is always 0.5$.

### Variable Improvement_surcharge
EDA_Univariant(df[,"Improvement_surcharge"],Improvement_surcharge)

# errors
l <- which(df$Improvement_surcharge<0)
if (length(l)>0) {
  ierrs[l]<-ierrs[l]+1
  jerrs["Improvement_surcharge"]<-length(l)
}
df[l,"Improvement_surcharge"]<-NA 
df[,"Improvement_surcharge"] <- as.factor(df[,"Improvement_surcharge"] == 0.3)
EDA_Univariant(df[,"Improvement_surcharge"],Improvement_surcharge)

#it is not a useful information because in theory it is always 0.3$.

# Variable Tip_amount
hist(df$Tip_amount, main="Histogram of Tip_amount")
boxplot(df$Tip_amount, main="Boxplot of Tip_amount")
summary(df$Tip_amount)

# errors
l <- which(df$Tip_amount<0)
if (length(l)>0) {
  ierrs[l]<-ierrs[l]+1
  jerrs["Tip_amount"]<-length(l)
}
df[l,"Tip_amount"]<-NA 

# Outlier detection
boxplot(df$Tip_amount)
var_out<-calcQ(df$Tip_amount)
abline(h=var_out$souts,col="red")
abline(h=var_out$souti,col="blue")
var_out$souts
l <-which(df$Tip_amount>20)
iouts[l ]<-iouts[l]+1
jouts["Tip_amount"]<-length(l)
df[l,"Tip_amount"]<-NA 

hist(df$Tip_amount, main="Histogram of Tip_amount")
boxplot(df$Tip_amount, main="Boxplot of Tip_amount")
summary(df$Tip_amount)


### Variable Tolls_amount

hist(df$Tolls_amount, main="Histogram of Tolls_amount")
boxplot(df$Tolls_amount, main="Boxplot of Tolls_amount")
summary(df$Tolls_amount)

# errors
l <- which(df$Tolls_amount<0)
if (length(l)>0) {
  ierrs[l]<-ierrs[l]+1
  jerrs["Tolls_amount"]<-length(l)
}
df[l,"Tolls_amount"]<-NA 

# Outlier detection
boxplot(df$Tolls_amount)
var_out<-calcQ(df$Tolls_amount)
abline(h=var_out$souts,col="red")
abline(h=var_out$souti,col="blue")
var_out$souts
l <-which(df$Tolls_amount>7)
iouts[l ]<-iouts[l]+1
jouts["Tolls_amount"]<-length(l)
df[l,"Tolls_amount"]<-NA 

hist(df$Tolls_amount, main="Histogram of Tolls_amount")
boxplot(df$Tolls_amount, main="Boxplot of Tolls_amount")
summary(df$Tolls_amount)


### Variable Total_amount

#errors

summary(df$Total_amount)
l<-which(df$Total_amount<0)
if (length(l)>0) {
  ierrs[l]<-ierrs[l]+1
  jerrs["Total_amount"]<-length(l)
}
df[l,"Total_amount"]<-NA

# Outlier detection
boxplot(df$Total_amount)
var_out<-calcQ(df$Total_amount)
abline(h=var_out$souts,col="red")
abline(h=var_out$souti,col="blue")
l<-which((df$Total_amount<0)|(df$Total_amount>100))
iouts[l]<-iouts[l]+1
jouts["Total_amount"]<-length(l)
df[l,"Total_amount"]<-NA 


### Variable Ehail_fee
summary(df$Ehail_fee)
df$Ehail_fee<-NULL

```





## Define new variables

Create new variables derived from the original ones, as effective speed, travel time, hour of request, period of request, effective trip distance (in km) 

```{r}
# Trip length in km
df$tlenkm<-df$Trip_distance*1.609344 # Miles to km
summary(df$tlenkm)

# Travel time in min
df$traveltime<-(as.numeric(as.POSIXct(df$lpep_dropoff_datetime)) - as.numeric(as.POSIXct(df$lpep_pickup_datetime)))/60

#errors
summary(df$traveltime)
l<-which(df$traveltime<0);length(l)
if (length(l)>0) {
  ierrs[l]<-ierrs[l]+3
  jerrs["traveltime"]<-length(l)
  jerrs["lpep_pickup_datetime"] <- length(l)
  jerrs["lpep_dropoff_datetime"] <- length(l)
}
df[l,"traveltime"]<-NA 
df[l,"lpep_dropoff_datetime"] <- NA
df[l,"lpep_pickup_datetime"] <- NA

#outliers
boxplot(df$traveltime)
var_out<-calcQ(df$traveltime)
abline(h=var_out$souts,col="red")
abline(h=var_out$souti,col="blue")
l<-which((df$traveltime<0)|(df$traveltime>800))
iouts[l]<-iouts[l]+1
jouts["traveltime"]<-length(l)
df[l,"traveltime"]<-NA 
df[l,"lpep_dropoff_datetime"] <- NA
df[l,"lpep_pickup_datetime"] <- NA


# Effective speed (km/h)
df$espeed<-(df$tlenkm/(df$traveltime))*60
summary(df$espeed)


# errors
summary(df$espeed)
l<-which((df$espeed<=0)|(df$espeed=="Inf"))
ierrs[l]<-ierrs[l]+1
jerrs["espeed"]<-length(l)
df[l,"espeed"]<-NA 

# outliers
summary(df$espeed)
Boxplot(df$espeed)
var_out<-calcQ(df$espeed)
abline(h=var_out$souts,col="red")
abline(h=var_out$souti,col="blue")

l<-which((df$espeed<=3)|(df$espeed>240))
iouts[l]<-iouts[l]+1
jouts["espeed"]<-length(l)
df[l,"espeed"]<-NA 



df$lpep_pickup_time<-as.numeric(substr(strptime(df$lpep_pickup_datetime, "%Y-%m-%d %H:%M:%S"),12,13))
df$lpep_pickup_period<-1
df$lpep_pickup_period[df$lpep_pickup_time>7]<-2
df$lpep_pickup_period[df$lpep_pickup_time>10]<-3
df$lpep_pickup_period[df$lpep_pickup_time>16]<-4
df$lpep_pickup_period[df$lpep_pickup_time>19]<-1
df$lpep_pickup_period<-factor(df$lpep_pickup_period,labels=paste("Period",c("night","morning","valley","afternoon")))
df$lpep_dropoff_time<-as.numeric(substr(strptime(df$lpep_dropoff_datetime, "%Y-%m-%d %H:%M:%S"),12,13))
df$lpep_dropoff_period<-1
df$lpep_dropoff_period[df$lpep_dropoff_time>7]<-2
df$lpep_dropoff_period[df$lpep_dropoff_time>10]<-3
df$lpep_dropoff_period[df$lpep_dropoff_time>16]<-4
df$lpep_dropoff_period[df$lpep_dropoff_time>19]<-1
df$lpep_dropoff_period<-factor(df$lpep_dropoff_period,labels=paste("Period",c("night","morning","valley","afternoon")))

### Variable distHaversine
library(geosphere) 
df$distHaversine <-distHaversine(df[,c("Pickup_longitude", "Pickup_latitude")],
              df[,c("Dropoff_longitude", "Dropoff_latitude")])/1000
# Errors
summary(df$distHaversine)
l<-which((df$distHaversine<0)|(df$distHaversine=="Inf"))
ierrs[l]<-ierrs[l]+1
jerrs["distHaversine"]<-length(l)
df[l,"distHaversine"]<-NA 




## Variable lpep_pickup_date, 

df_datatime <- t(as.data.frame(strsplit(as.character(df$lpep_pickup_datetime), " ")))
df$lpep_pickup_date <- factor(df_datatime[,1])


```



## Imputation of qualitative variables

```{r}

# vars_dis <- c( "lpep_pickup_datetime", "lpep_dropoff_datetime", "MTA_tax", "Improvement_surcharge")
# 
# summary(df[,vars_dis])
# res.immca<-imputeMCA(df[,vars_dis],ncp=10)
# summary(res.immca$completeObs)
# 
# # Check one by one
# df[ , vars_dis]<-res.immca$completeObs  

```


## Imputation of numeric variables

```{r}
names(df)
summary(df)
#vars_con
vars_con <- c("Pickup_longitude", "Pickup_latitude", "Dropoff_longitude", "Dropoff_latitude","Fare_amount", "Tip_amount", "Tolls_amount", "Total_amount", "espeed", "traveltime", "tlenkm", "distHaversine")

summary(df[,vars_con])
res.impca<-imputePCA(df[,vars_con],ncp=6)
summary(res.impca$completeObs)
df[ ,vars_con]<-res.impca$completeObs
summary(df[,vars_con])

# Check one by one

#espeed
l<-which(df[,"espeed"]<0)
df[l,"espeed"]<-0.001

#Trip_distance
l<-which(df[,"Trip_distance"]<0)
df[l,"Trip_distance"]<-0.001

#traveltime
l<-which(df[,"traveltime"]<0)
df[l,"traveltime"]<-0.001

#tlenkm
l<-which(df[,"tlenkm"]<0)
df[l,"tlenkm"]<-0.001

#distHaversine
l<-which(df[,"distHaversine"]<0)
df[l,"distHaversine"]<-0.001


summary(df)
```

## Multivariant Outliers

```{r}
vars_con <- c( "tlenkm","Fare_amount", "Total_amount")
summary(df[,vars_con])
aq.plot(df[,vars_con],delta=qchisq(0.995,length(vars_con)),quan=0.75)
mout<-Moutlier(df[,vars_con],quantile = 0.995, plot = TRUE)
par(mfrow=c(1,1))
plot(mout$md,mout$rd, type="n")
text(mout$md,mout$rd,labels=vars_con)

Boxplot(mout$rd)
summary(mout$rd)
l<-which(mout$rd>50);length(l)
df[,"multiouts"] <- FALSE
df[l,"multiouts"] <- TRUE
df[,"multiouts"] <- as.factor(df[l,"multiouts"])
```

## Discretization 

```{r}
## Variable espeed

summary(df[,vars_con])

varaux<-factor(cut(df$espeed,breaks=quantile(df$espeed,seq(0,1,0.25),na.rm=TRUE),include.lowest = T ))
summary(varaux)
tapply(df$espeed,varaux,median)  #tapply(X, INDEX, FUN = NULL) map function
df$f.espeed<-factor(cut(df$espeed,breaks=c(0,25,max(df$espeed),na.rm=TRUE),include.lowest = T ))
levels(df$f.espeed)<-paste("f.espeed-",levels(df$f.espeed),sep="")

## Variable tlenkm
summary(df$tlenkm)
varaux<-factor(cut(df$tlenkm,breaks=quantile(df$tlenkm,seq(0,1,0.25),na.rm=TRUE),include.lowest = T ))
summary(varaux)
df$f.tlenkm<-factor(cut(df$tlenkm,breaks=c(0,5,max(df$tlenkm),na.rm=TRUE),include.lowest = T ))
levels(df$f.tlenkm)<-paste("f.tlenkm-",levels(df$f.tlenkm),sep="")

## Variable traveltime
summary(df$traveltime)
varaux<-factor(cut(df$traveltime,breaks=c(0,15,40,max(df$traveltime)),include.lowest = T ))
summary(varaux)
df$f.traveltime <- varaux
levels(df$f.traveltime)<-paste("f.traveltime-",levels(df$f.traveltime),sep="")

## Variable distHaversine

varaux<-factor(cut(df$distHaversine,breaks=quantile(df$distHaversine,seq(0,1,0.25),na.rm=TRUE),include.lowest = T ))
summary(varaux)
tapply(df$distHaversine,varaux,median)  #tapply(X, INDEX, FUN = NULL) map function
df$f.distHaversine<-factor(cut(df$distHaversine,breaks=c(0,5,10,max(df$distHaversine)),include.lowest = T ))
levels(df$f.distHaversine)<-paste("f.distHaversine-",levels(df$f.distHaversine),sep="")

# Variable AnyToll
Boxplot(df$Tolls_amount)
varaux<-factor(cut(df$Tolls_amount,breaks=c(0,0.01,max(df$Tolls_amount)),include.lowest = T ))
df$AnyToll <- df$Tolls_amount < 0.01
df$AnyToll<-factor(df$AnyToll,labels=paste("AnyToll",c("No","Yes")))
summary(df$AnyToll)  


## Variable AnyTip

df$AnyTip<-ifelse(df$Tip_amount<0.01,0,1)
df$AnyTip<-factor(df$AnyTip,labels=paste("AnyTip",c("No","Yes")))


```

## Data Quality report (missings/errors/outliers)

```{r}

#Variable missings
EDA_Univariant(jmis,"Missings")
pie(jmis, labels=names(jmis),main = "Pie jmis")
jmis["Ehail_fee"] 

### Variable errors
l <-which(ierrs>1)
pie(ierrs[l], labels = 1:length(l))
l <-which(ierrs>2)
pie(ierrs[l], labels = 1:length(l))
l <-which(ierrs>3)
pie(ierrs[l], labels = 1:length(l))
summary(jerrs)
pie(jerrs, labels=names(jerrs),main = "Pie jerrs")

### Variable outliers
l <-which(iouts>1)
pie(iouts[l], labels = 1:length(l))
l <-which(iouts>2)
pie(iouts[l], labels = 1:length(l))
summary(jouts)
pie(jouts, labels=names(jouts),main = "Pie jouts")
```

## Univariant Exploratory Analysis (EDA)

```{r}
summary(df)
vars_dis <- c("VendorID", "RateCodeID","Payment_type", "Store_and_fwd_flag", "lpep_pickup_period", "lpep_dropoff_period", "Trip_type")

vars_con <- c( "Passenger_count", "Trip_distance", "Pickup_longitude", "Pickup_latitude", "Dropoff_longitude", "Dropoff_latitude","Fare_amount", "Extra", "MTA_tax", "Improvement_surcharge", "Tip_amount", "Tolls_amount", "Total_amount")


#VendorID
EDA_Univariant(df[,"VendorID"],"VendorID")

#lpep_pickup_period, lpep_pick_hour, traveltime, f.traveltime
EDA_Univariant(df[,c("lpep_pickup_time", "lpep_pickup_period","traveltime","f.traveltime")])


#tlenkm
EDA_Univariant(df[,c("tlenkm","f.tlenkm")])

#espeed

EDA_Univariant(df[,c("espeed","f.espeed")])


#RateCodeID
EDA_Univariant(df[,"RateCodeID"],"RateCodeID")

#Points
plot(df$Dropoff_longitude, df$Dropoff_latitude, pch = 20, col = "blue" , xlab="longitude", ylab="latitudel")
points(df$Pickup_longitude, df$Pickup_latitude, pch = 20, col = "red")
legend("topleft", legend=c("Dropoff", "Pickup"),
       col=c("blue", "red"), lty=1)



#Store_and_fwd_flag
EDA_Univariant(df[,c("Store_and_fwd_flag")], "Store_and_fwd_flag")

#Passenger_count
EDA_Univariant(df[,c("Passenger_count")], "Passenger_count")

#Payment_type
EDA_Univariant(df[,c("Payment_type")], "Payment_type")


#Extra, MTA_tax, Improvement_surcharge
EDA_Univariant(df[,c("f.Extra", "MTA_tax", "Improvement_surcharge")])


#Fare_amount, Tip_amount, Tolls_amount, Total_amount
EDA_Univariant(df[,c("Fare_amount", "Tip_amount", "Tolls_amount", "Total_amount")])
EDA_Univariant(df[,c("AnyTip", "AnyToll")])

#Trip_type
EDA_Univariant(df[,c("Trip_type")])

vars_dis <- c("VendorID", "RateCodeID","Payment_type", "Store_and_fwd_flag", "lpep_pickup_period", "lpep_dropoff_period", "Trip_type", "lpep_pickup_period", "lpep_dropoff_period")
vars_con <- c( "Passenger_count", "Trip_distance", "Pickup_longitude", "Pickup_latitude", "Dropoff_longitude", "Dropoff_latitude","Fare_amount", "Extra", "MTA_tax", "Improvement_surcharge", "Tip_amount", "Tolls_amount", "Total_amount")

```


## Profiling

```{r}

# Total_amount
i <- names(df)
i <- which(i=="Total_amount")
res.condes<-condes(df,i)  
res.condes$quanti 
res.condes$quali 
res.condes$category  
#AnyTip
i <- names(df)
i <- which(i=="AnyTip")
res.cat<-catdes(df,i)
res.cat$quanti.var
res.cat$quanti
res.cat$category



```


